---
title: "Case Study W01"
author: "Carlos Gerez"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:  
    keep_md: true
    toc: true
    toc_float: true
    code_folding: hide
    fig_height: 6
    fig_width: 12
    fig_align: 'center'
---

```{r, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r load_libraries, include=FALSE}
# Use this R-Chunk to load all your libraries!

#install.packages("tidyverse") # run this line once in console to get package
#library(tidyverse)

```

```{r load_data}
# Use this R-Chunk to import all your datasets!

```

## Background

 Visualization Critique of three article found in the web.
 
 This are the links to the three articles: 
 
 1. [Are Pop Lyrics Getting More Repetitive?](https://pudding.cool/2017/05/song-repetition/)
 
 2. [Club Soccer Predictions](https://projects.fivethirtyeight.com/soccer-predictions/?ex_cid=rrpromo)
 
 3. [How many people die and how many are born each year?](https://ourworldindata.org/births-and-deaths)
 
 



```{r tidy_data}
# Use this R-Chunk to clean & wrangle your data!


```

## Data Visualization

```{r plot_data}
# Use this R-Chunk to plot & visualize your data!
plot(x = iris$Sepal.Length,
     y = iris$Petal.Length, 
     col = iris$Species, pch = 19)
```

## Conclusions
  1. **Are Pop Lyrics Getting More Repetitive?**  
    1.1 What is unique or good about the visualization? List a few ideas.  
      In this article the repetitiveness of song lyrics is measure is measure using [Lempel-Ziv algorithm](https://en.wikipedia.org/wiki/LZ77_and_LZ78).The first part is a great visual explanation with text samples of how the algorithm work and is used to obtain the data, and in the second part the results are shown with visual graphs.  
      Good points in this visual presentation are:
      <br/><br/>
        a. A good visual explanation on the pre-processing that is key to understand the results.  
        b. The data distribution is shown in a kind of standardized form (using logarithmic scale) and in the real form to see for differences in the appreciations.  
        c. In the final analysis the variation in the repetitiveness is clearly shown in percent by year in a good example of discrimination and emergence.  
        d. In the conclusions ("who is responsible for this madness") helps to understand by year, artist and discography the repetitiveness in songs development by different levels and add new insights about repetitiveness in the top tens of the charts.  
        
     1.2 What could be better? Identify at least one issue with the visualization.
        <br/><br/>
        a. It could help to have an automatic slider to show the progression by year and artist in the last graph of repetitiveness  by artist.  
        b. In the same graph it should be a better way to get the circles size  proportional to the average repetition.  
        c. Almost all the graph need some kind of explanation to be understand, then there is not a emergent (gestalt) componet obvious to the viewer.  
        
        
***
   2. **Club Soccer Predictions**  
      2.1 What is unique or good about the visualization? List a few ideas.  
      This visualization shows probabilities of rankings in the top of different championships in a simple and direct way including ratings generated by an algorithm that they explain in the top link : [How this works](https://fivethirtyeight.com/methodology/how-our-club-soccer-predictions-work/).
      <br/><br/>
        a. In the opening page since they show the three more probable teams to reach the finals, they use tables instead of pie charts.  
        b. The use color saturation and hue helps to guide the viewer to evaluate the results (detection).
      
      2.2 What could be better? Identify at least one issue with the visualization
      <br/><br/>
        a. It will help in football to show the progression of the teams in a graph, since soccer are full of surprise last minutes goals, that can show also how the attitude of the teams reflects in the final outcomes. For example when a team looses several games in a row, or coming from winning two or more games in the extra time.  
        b. I had to take some time to understand how percent works in this context of probabilities when see for example higher percents in the second row of making to the finals than in wining it (the case of Juventus in the Europa league). 
      
     
      
       
***
   3. **How many people die and how many are born each year?**  
      3.1 What is unique or good about the visualization? List a few ideas.  
      This visualization of number of births and death across the world shows the numbers of births by continent and the world as a whole with the data we have today and prognosis for the future until 2100 from the United Nations World Population Projects.
      <br/><br/>
        a. It shows in a common scale both easy to compare the growth of both numbers.  
        b. It uses an area graph to show the increase comparative by continents.
      
      3.2 What could be better? Identify at least one issue with the visualization
      <br/><br/>
        a. In the area graphs is difficult to see the line between Oceania and Latin America.  
        b. It could help to infer future growth or decrease of the global actual and estimates shown in a graphical way. The percents and estimates are named in the conclusions but not shown.  
   
   
      

