---
title: "Interactive Maps"
author: "Carlos Gerez"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:  
    keep_md: true
    toc: true
    toc_float: true
    code_folding: hide
    fig_height: 6
    fig_width: 8
    fig_align: 'center'
---

```{r, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r load_libraries, include=FALSE}
# Use this R-Chunk to load all your libraries!
#install.packages("tidyverse") # run this line once in console to get package
#install.packages("xfun")
#install.packages("leaflet.extras2")
#install.packages("leafletCN")
library(tidyverse)
library(shiny)
library(raster)
library(shiny)
library(leaflet)
library( crosstalk )
library(RColorBrewer)
library(USAboundaries)
library(sf, warn.conflicts = F)
library(leaflet.extras2)
library(leafletCN)
```

```{r load_data}
# Use this R-Chunk to import all your datasets!
states <- us_states()
counties <- us_counties()
#names(counties)
counties <- counties %>% dplyr::select(-9)
buildings_p_state_t <- buildings::permits 



```

## Background

1. Choose an option:  
    - Recreate a previous task or case study visualization with leaflet OR  
    - Find your own spatial data that interests you and build a visualization  
2. Publish your case study to Rstudio Connect at shiny.byui.edu
## Data Wrangling

```{r tidy_data}

buildings_p_state <- buildings_p_state_t %>% group_by(StateAbbr,year, variable ) %>% 
  summarise(StateAbbr,
            permitTotals = sum(value)) %>% 
  distinct()



statesBuildings <-   buildings_p_state %>% 
  full_join(states, by = c("StateAbbr" = "state_abbr") )

################# THis is the base data for the states graphs ############

statesBuildings <- statesBuildings %>% filter(
  variable == "Single Family")  

##########################################################################

buildings_p_county <- buildings_p_state_t %>% group_by(countyname, year, variable) %>% 
  summarise(countyname,
            StateAbbr,
            permitTotal = sum(value)) %>% filter(
              variable == "Single Family")  


###################### This is the base data for the county maps ############

countyBuildings <-   buildings_p_county %>% 
  full_join(counties, by = c("countyname" = "namelsad") )

#############################################################################
```

## Data Visualization
### I recreate the data from single family permits from 1980 to 2010.
```{r plot_data}
# Use this R-Chunk to plot & visualize your data!


############### start with a selection of year and state #############
# range of years is 1980 to 2010
# the idea is show how getting in the future the information from a slider (shiny) this can be implemented

selected_year <- 1980
selected_state <- "Utah"

selectedState <- statesBuildings %>% filter(year == selected_year) %>% st_as_sf()

selected <- countyBuildings %>% filter(state_name == selected_state) %>% st_as_sf()
selected <- selected %>% filter(year == selected_year)


#  Create the palettes for each legends added

#pal <- brewer.pal(12, "Set3")
#statePal <- colorFactor(pal, selectedState$name)

# First Graph palette ###############################

pal2 <- colorNumeric(
  palette = "Blues",
  domain = selectedState$permitTotals / 1000
)

# Second graph palette #####################
countyPal <- colorNumeric(
        palette = "Blues"  ,
        domain = selected$permitTotal[selected$year == selected_year])

################# Create labels an titles formatted #####################
### first graph titles

labels <- sprintf(
  "<strong>%s</strong><br/>%g permits in year %i<sup></sup>",
  selectedState$name, selectedState$permitTotals, selectedState$year
) %>% lapply(htmltools::HTML)

titles <- sprintf(
  "<strong>Number of Permits</strong><br/>(1000)<br/> <strong>in Year </strong>%i<sup></sup>",
  selected_year
) %>% lapply(htmltools::HTML)

######### Second graph titles ######################

label2 <- sprintf(
  "<strong>%s</strong><br/><strong>%s</strong><br/>%g permits in year %i<sup></sup>",
  selected$countyname,selected$state_name, selected$permitTotal, selected$year
) %>% lapply(htmltools::HTML)


title2 <-  sprintf(
  "<strong>Number of Permits</strong><br/><strong>in Year </strong>%i<sup></sup>",
  selected_year
) %>% lapply(htmltools::HTML)
 
######################### first leaflet ############################ 

leaflet() %>% 
        addTiles() %>% 
        setView(lat = 37.58 , lng = -103.46, zoom = 3.5) %>% 
        addPolygons(data = selectedState, 
                    fillColor = ~pal2(selectedState$permitTotals / 1000), 
                    fillOpacity = 1, 
                    color = "black", 
                    stroke = T, 
                    weight = 1, 
#                   layerId = states$state_name, #this sets the click id, very important! when translated to shiny
                    highlightOptions = highlightOptions(color = "black",
                                                        weight = 3,
                                                        bringToFront = TRUE),
                    label = labels,
                    labelOptions = labelOptions(style = list("font-weight" = "normal",padding = "3px 8px"),
                                              textsize = "15px",
                                              direction = "auto")) %>% 
        addLegend(data = selectedState, values = ~selectedState$permitTotals / 1000,group = "1980", className = "info legend 1980", 
                  position =  "bottomleft",
                  title = titles,
                  labFormat = labelFormat(digits = 3, 
                  big.mark = ""), pal = pal2) 
   

#####################################################################################


################## Second leaflet ###########################################


leaflet() %>% 
            addTiles() %>% 
            addPolygons(data = selected, 
                        fillColor = ~countyPal(selected$permitTotal), 
                        fillOpacity = 1, 
                        color = "black", 
                        stroke = T, 
                        weight = 1, 
#                    layerId = states$state_name, #this sets the click id, very important!
                    highlightOptions = highlightOptions(color = "black",
                                                        weight = 3,
                                                        bringToFront = TRUE),
                    label = label2,
                    labelOptions = labelOptions(style = list("font-weight" = "normal",
                                              padding = "3px 8px"),
                                              textsize = "15px",
                                              direction = "auto")) %>% 
                      addLegend(data = selected, values = ~selected$permitTotal,group = "1980", className = "info legend 1980", 
                                position =  "bottomleft",
                                title = title2,
                                labFormat = labelFormat(big.mark = ""), pal = countyPal)
                        


################## Working with groups for more interactivity ##################


######## create diferent sets to display with controls ################
################### must find a way in the future to do it with a function ( a list probably) #########

selected_year1980 <- 1980
selected_year1985 <- 1985
selected_year1990 <- 1990
selected_year1995 <- 1995
selected_year2000 <- 2000
selected_year2005 <- 2005
selected_year2010 <- 2010



#selected_state <- "Utah"

selectedState1980 <- statesBuildings %>% filter(year == selected_year1980) %>% st_as_sf()
selectedState1985 <- statesBuildings %>% filter(year == selected_year1985) %>% st_as_sf()
selectedState1990 <- statesBuildings %>% filter(year == selected_year1990) %>% st_as_sf()
selectedState1995 <- statesBuildings %>% filter(year == selected_year1995) %>% st_as_sf()
selectedState2000 <- statesBuildings %>% filter(year == selected_year2000) %>% st_as_sf()
selectedState2005 <- statesBuildings %>% filter(year == selected_year2005) %>% st_as_sf()
selectedState2010 <- statesBuildings %>% filter(year == selected_year2010) %>% st_as_sf()





#selected <- countyBuildings %>% filter(state_name == selected_state) %>% st_as_sf()
#selected <- selected %>% filter(year == selected_year)


################ palettes for each year, latter can be simplify ###############


# a function to create each palette function

generate_palette <- function(selected){
  pal2 <- colorNumeric(
  palette = "Blues",
  domain = selected$permitTotals / 1000
)
}

pal1980 <- generate_palette(selectedState1980)

pal1985 <- generate_palette(selectedState1985)

pal1990 <- generate_palette(selectedState1990)

pal1995 <- generate_palette(selectedState1995)

pal2000 <- generate_palette(selectedState2000)

pal2005 <- generate_palette(selectedState2005)

pal2010 <- generate_palette(selectedState2010)

############################### Create labels and titles for each group ###################
# pending a function to do this work with sufix, prefix and or paste0

####################### 1980

labels1980 <- sprintf(
  "<strong>%s</strong><br/>%g permits in year %i<sup></sup>",
  selectedState1980$name, selectedState1980$permitTotals, selectedState1980$year
) %>% lapply(htmltools::HTML)

titles1980 <- sprintf(
  "<strong>Number of Permits</strong><br/>(1000)<br/> <strong>in Year </strong>%i<sup></sup>",
  selected_year1980
) %>% lapply(htmltools::HTML)

################### 1985

labels1985 <- sprintf(
  "<strong>%s</strong><br/>%g permits in year %i<sup></sup>",
  selectedState1985$name, selectedState1985$permitTotals, selectedState1985$year
) %>% lapply(htmltools::HTML)

titles1985 <- sprintf(
  "<strong>Number of Permits</strong><br/>(1000)<br/> <strong>in Year </strong>%i<sup></sup>",
  selected_year1985
) %>% lapply(htmltools::HTML)

#################### 1990

labels1990 <- sprintf(
  "<strong>%s</strong><br/>%g permits in year %i<sup></sup>",
  selectedState1990$name, selectedState1990$permitTotals, selectedState1990$year
) %>% lapply(htmltools::HTML)

titles1990 <- sprintf(
  "<strong>Number of Permits</strong><br/>(1000)<br/> <strong>in Year </strong>%i<sup></sup>",
  selected_year1990
) %>% lapply(htmltools::HTML)

##################### 1995

labels1995 <- sprintf(
  "<strong>%s</strong><br/>%g permits in year %i<sup></sup>",
  selectedState1995$name, selectedState1995$permitTotals, selectedState1995$year
) %>% lapply(htmltools::HTML)

titles1995 <- sprintf(
  "<strong>Number of Permits</strong><br/>(1000)<br/> <strong>in Year </strong>%i<sup></sup>",
  selected_year1995
) %>% lapply(htmltools::HTML)


######################### 2000

labels2000 <- sprintf(
  "<strong>%s</strong><br/>%g permits in year %i<sup></sup>",
  selectedState2000$name, selectedState2000$permitTotals, selectedState2000$year
) %>% lapply(htmltools::HTML)

titles2000 <- sprintf(
  "<strong>Number of Permits</strong><br/>(1000)<br/> <strong>in Year </strong>%i<sup></sup>",
  selected_year2000
) %>% lapply(htmltools::HTML)

########################## 2005

labels2005 <- sprintf(
  "<strong>%s</strong><br/>%g permits in year %i<sup></sup>",
  selectedState2005$name, selectedState2005$permitTotals, selectedState2005$year
) %>% lapply(htmltools::HTML)

titles2005 <- sprintf(
  "<strong>Number of Permits</strong><br/>(1000)<br/> <strong>in Year </strong>%i<sup></sup>",
  selected_year2005
) %>% lapply(htmltools::HTML)

########################## 2010

labels2010 <- sprintf(
  "<strong>%s</strong><br/>%g permits in year %i<sup></sup>",
  selectedState2010$name, selectedState2010$permitTotals, selectedState2010$year
) %>% lapply(htmltools::HTML)

titles2010 <- sprintf(
  "<strong>Number of Permits</strong><br/>(1000)<br/> <strong>in Year </strong>%i<sup></sup>",
  selected_year2010
) %>% lapply(htmltools::HTML)







############################### create graphs with groups and controls ##############

p1 <- leaflet() %>% 
        addTiles() %>% 
        setView(lat = 37.58 , lng = -103.46, zoom = 3.5) %>% 
        addPolygons(data = selectedState1980, group = "1980",
                    fillColor = ~pal1980(selectedState1980$permitTotals / 1000), 
                    fillOpacity = 1, 
                    color = "black", 
                    stroke = T, 
                    weight = 1,
                    highlightOptions = highlightOptions(color = "black",
                                                        weight = 3,
                                                        bringToFront = TRUE),
                    label = labels1980,
                    labelOptions = labelOptions(style = list("font-weight" = "normal",padding = "3px 8px"),
                                              textsize = "15px",
                                              direction = "auto")) %>% 
        addLegend(data = selectedState1980, group = "1980", className = "info legend 1980", values = ~selectedState1980$permitTotals / 1000,
                  position =  "bottomleft",
                  title = titles1980,
                  labFormat = labelFormat(digits = 3, 
                  big.mark = ""), pal = pal1980) %>%
  ################ 1985
        addPolygons(data = selectedState1985, group = "1985", 
                    fillColor = ~pal1985(selectedState1985$permitTotals / 1000), 
                    fillOpacity = 1, 
                    color = "black", 
                    stroke = T, 
                    weight = 1,
                    highlightOptions = highlightOptions(color = "black",
                                                        weight = 3,
                                                        bringToFront = TRUE),
                    label = labels1985,
                    labelOptions = labelOptions(style = list("font-weight" = "normal",padding = "3px 8px"),
                                              textsize = "15px",
                                              direction = "auto")) %>% 
        addLegend(data = selectedState1985, group = "1985", className = "info legend 1985", values = ~selectedState1985$permitTotals / 1000,
                  position =  "bottomleft",
                  title = titles1985,
                  labFormat = labelFormat(digits = 3, 
                  big.mark = ""), pal = pal1985) %>% 
  ################## 1990
        addPolygons(data = selectedState1990, group = "1990", 
                    fillColor = ~pal1990(selectedState1990$permitTotals / 1000), 
                    fillOpacity = 1, 
                    color = "black", 
                    stroke = T, 
                    weight = 1,
                    highlightOptions = highlightOptions(color = "black",
                                                        weight = 3,
                                                        bringToFront = TRUE),
                    label = labels1990,
                    labelOptions = labelOptions(style = list("font-weight" = "normal",padding = "3px 8px"),
                                              textsize = "15px",
                                              direction = "auto")) %>% 
        addLegend(data = selectedState1990, group = "1990",  className = "info legend 1990",values = ~selectedState1990$permitTotals / 1000,
                  position =  "bottomleft",
                  title = titles1990,
                  labFormat = labelFormat(digits = 3, 
                  big.mark = ""), pal = pal1990) %>% 
  ########### 1995
        addPolygons(data = selectedState1995, group = "1995", 
                    fillColor = ~pal1995(selectedState1995$permitTotals / 1000), 
                    fillOpacity = 1, 
                    color = "black", 
                    stroke = T, 
                    weight = 1,
                    highlightOptions = highlightOptions(color = "black",
                                                        weight = 3,
                                                        bringToFront = TRUE),
                    label = labels1995,
                    labelOptions = labelOptions(style = list("font-weight" = "normal",padding = "3px 8px"),
                                              textsize = "15px",
                                              direction = "auto")) %>% 
        addLegend(data = selectedState1995, group = "1995", className = "info legend 1995", values = ~selectedState1995$permitTotals / 1000,
                  position =  "bottomleft",
                  title = titles1995,
                  labFormat = labelFormat(digits = 3, 
                  big.mark = ""), pal = pal1995) %>% 
  ############# 2000
        addPolygons(data = selectedState2000, group = "2000", 
                    fillColor = ~pal2000(selectedState2000$permitTotals / 1000), 
                    fillOpacity = 1, 
                    color = "black", 
                    stroke = T, 
                    weight = 1,
                    highlightOptions = highlightOptions(color = "black",
                                                        weight = 3,
                                                        bringToFront = TRUE),
                    label = labels2000,
                    labelOptions = labelOptions(style = list("font-weight" = "normal",padding = "3px 8px"),
                                              textsize = "15px",
                                              direction = "auto")) %>% 
        addLegend(data = selectedState2000, group = "2000", className = "info legend 2000", values = ~selectedState2000$permitTotals / 1000,
                  position =  "bottomleft",
                  title = titles2000,
                  labFormat = labelFormat(digits = 3, 
                  big.mark = ""), pal = pal2000)  %>% 
  ############## 2005
        addPolygons(data = selectedState2005, group = "2005", 
                    fillColor = ~pal2005(selectedState2005$permitTotals / 1000), 
                    fillOpacity = 1, 
                    color = "black", 
                    stroke = T, 
                    weight = 1,
                    highlightOptions = highlightOptions(color = "black",
                                                        weight = 3,
                                                        bringToFront = TRUE),
                    label = labels2005,
                    labelOptions = labelOptions(style = list("font-weight" = "normal",padding = "3px 8px"),
                                              textsize = "15px",
                                              direction = "auto")) %>% 
        addLegend(data = selectedState2005, group = "2005", className = "info legend 2005", values = ~selectedState2005$permitTotals / 1000,
                  position =  "bottomleft",
                  title = titles2005,
                  labFormat = labelFormat(digits = 3, 
                  big.mark = ""), pal = pal2005) %>% 
  ############### 2010
  
        addPolygons(data = selectedState2010, group = "2010", 
                    fillColor = ~pal2010(selectedState2010$permitTotals / 1000), 
                    fillOpacity = 1, 
                    color = "black", 
                    stroke = T, 
                    weight = 1,
                    highlightOptions = highlightOptions(color = "black",
                                                        weight = 3,
                                                        bringToFront = TRUE),
                    label = labels2010,
                    labelOptions = labelOptions(style = list("font-weight" = "normal",padding = "3px 8px"),
                                              textsize = "15px",
                                              direction = "auto")) %>% 
        addLegend(data = selectedState2010, group = "2010",  className = "info legend 2010",values = ~selectedState2010$permitTotals / 1000,
                  position =  "bottomleft",
                  title = titles2010,
                  labFormat = labelFormat(digits = 3, 
                  big.mark = ""), pal = pal2010) %>% 
  ############## Layers controls of the groups
        addLayersControl(
    baseGroups = c("1980","1985","1990","1995","2000","2005","2010"),
#    overlayGroups = c("Quakes", "Outline"),
    options = layersControlOptions(collapsed = FALSE)
  ) %>% htmlwidgets::onRender("
      function(el, x) {
         var updateLegend = function () {
            var selectedGroup = document.querySelectorAll('input:checked')[0].nextSibling.innerText.substr(1);

            document.querySelectorAll('.legend').forEach(a => a.hidden=true);
            document.querySelectorAll('.legend').forEach(l => {
               if (l.classList.contains(selectedGroup)) l.hidden=false;
            });
         };
         updateLegend();
         this.on('baselayerchange', el => updateLegend());
      }"
   )

p1
```

## Conclusions  


I found that is more convenient for interactivity to wrap this leaftlet maps in shiny apps that offers a wider range of options and better output. For reasons of time that will be done in a future exploration.  
One of the discoveries is that some group implementations don't work yet for addLegend. I found a workaround on the internet that works well in this case. The problem is that also hide the first 2 graphs legends when submitted to the shiny server. I fixed that adding a group and a class that is the default in the last graph. That create legends that appears just when 1980 is selected.  
The 4 strategies for displaying map data help me to order my ideas of how to show the patterns of time in this case. Just to recall if need it:  
      1. Display data that is important  
      2. Aggregate the data into equal-sized hexagons (hexbins) to visualize patterns  
      3. Use scale visibility and/or multiscale rendering to show information when it is visually helpful.  
      4. Use clustering to group points into one symbol and apply Smart Mapping Styles  
wish I could use a slider to map the time, but I don’t find that in leaflet, just in shiny.  
I also thought that mapping the colors in time with a common scale will show better the patterns along the years in the single family permits.  

Finally in the first 2 graphs I choose a random year and state. The implementation was done thinking on how to get that information in the future from the user, for example choosing the state clicking in the USA map on a single state, and the years connected by a slider.

