---
title: "Week 4 tidyr"
output: 
  html_document:
    theme: cerulean
    code_folding: show
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE)
```

Take time to answer each question and create working code that you understand.

As you go through the activity, try to complete the exercise on your own. Each exercise is meant to take 5-10 minutes. If you are stuck or lost, try watching the code solution video linked at the bottom of the file.

## Background

Each semester BYU-Idaho holds a Research and Creative Works Conference (R&CW) where on-campus students can display key projects they have worked on during the semester. We will be analyzing the R&CW. Our goal is to see how many projects come from which departments over time.

## Read in data

To start, run this code to read in the data:


```{r echo=TRUE}
library(tidyverse, warn.conflicts = FALSE)
dat <- read_csv("https://byuistats.github.io/M335/data/rcw_reshape.csv")


dat


```

*Is the data in long or wide format? How can you tell?*

## Pivoting

Instructions:
1. Using what you learned in the reading, convert the data to wide format (one column per Department) and save it under a new name
2. Now convert the wide dataset back to long format

```{r wide_v_long}
#convert it to wide here
dat_wide <- dat %>% pivot_wider(names_from = Department, values_from = Count)
dat_wide

#now convert the wide data back to long format

dat_long <- dat_wide %>% pivot_longer(c(`Chem`,`CSEE`,`DCM`,`GEO`,`Math`,`ME`,`PHY`),names_to = "Department", values_to = "count")

dat_long <- dat_wide %>% pivot_longer(cols = Chem:PHY,names_to = "Department", values_to = "count")

dat_long

dat1 <- dat %>% pivot_wider(names_from = Department, values_from = Count) 
   
dat1 <- dat1 %>% pivot_longer(cols = Chem:PHY,names_to = "Department", values_to = "count")
```


### Discussion/Observation:

* *How many rows does the new long format have? *
* *How does that compare with the original data? Why?*
* *What argument in pivot_longer() could you change so that the "new" long format data matches the data you originally read in.*
*   *Discuss: When might you want to make implicit NA's explicit? Why would you not?*

## Diving deeper into implicit/explicit missing values

Instructions:

3. Explore the following related functions/arguments, make notes as desired
  - within pivot_wider( value_fill = )
  - within pivot_longer( values_drop_na = )
  - complete()
  - expand()
  
4. Using one (or more) of these commands, edit the original dataset so that the implicit missing values are made explicit; and instead of a missing value for "Count" they have zeros.  Save this dataset as zeros.

```{r zeros}
dat2 <- dat1

dat2 <- dat2 %>% complete()

dat2$count <- replace(dat2$count, is.na(dat2$count), 0)

dat2

dat2[is.na(dat2)] <- 0
dat2

#using pivot with values_fill
zeros <- dat %>% pivot_wider(names_from = Department, values_from = Count, values_fill = 0) %>%
  pivot_longer(cols = Chem:PHY, names_to = "Department", values_to = "Count")
zeros


# Using complete
# Just to show that add new combinations (emties)
dat %>% complete(Year,Semester, Department, fill = list(Count = 0, Semester_Date = "2") )

# work well without adding any new combination, and fill 
dat %>% complete(Department, nesting(Year, Semester, Semester_Date), fill = list(Count = 0))


# Expand with a left join to add the counts and replace-na
dat %>% expand(Department,nesting(Year, Semester, Semester_Date)) %>% 
  left_join(dat) %>% replace_na(list(Count = 0))

```



## Practice separate() and unite()

Starting with your zeros data:

5. Take Semester_Date column and separate it into a year, month, and day columns.
6. unite() the Year and Semester columns for potentially easier labelling of the graphic. (I actually prefer using paste() instead of unite(), so you could get familiar with that command as well)
7. Now create a line chart that shows the combined Year-Semester variable along the x axis and the count of projects on the y axis for each department.

```{r separate}
#5
# usse separate and change the type of the charachter day to numeric
dat2 <- dat2 %>% separate(col = Semester_Date, into = (c("day","month", "year"))) %>%
  mutate(day = as.numeric(day))



#6

# with paste into mutate

zeros %>% mutate(yrsem = paste(Year, Semester, sep = ""))


# with unite
zeros <- zeros %>% unite("yrsem",c(Year, Semester,), sep = "", remove = FALSE)

#dat2 <- dat2 %>% unite(new, month, year)
dat2 <- dat2 %>% unite(year_semester, Year, Semester, sep = "-")

zeros


#7 

ggplot(dat2) + 
  aes(x = year_semester, y =count, group = Department, color = Department) + 
  geom_line() +
#  facet_grid( . ~ year) +
 
  
   labs(x= "Year / Semester",
       y = "Count")

ggplot(zeros) + 
  aes(x = yrsem, y = Count, group = Department, color = Department) + 
  geom_line() +
#  facet_grid( . ~ year) +
 
 
   labs(x= "Year / Semester",
       y = "Count")

```

You may notice that the semesters are out of chronological order. That is because the new combined column of year and semester is probably being treated as a character vector, which will always be in alphabetical order.

We want to make the Year-Semester column a factor so that we can control/specify the order of the levels.

```{r}
#dat2 <- factor(dat2, "year_semester")
# factor year semester 

dat2 <- dat2 %>% mutate(year_semester = factor( level = c("2016Spring","2016Fall","2017Winter","2017Spring","2017Fall","2018Winter")))

ggplot(dat2) + 
  aes(x = year_semester, y = count, group = Department, color = Department) + 
  geom_line() +
#  facet_grid( . ~ year) +
 
  theme_bw() +
   labs(x= "Year / Semester",
       y = "Count")

```

[Part 1 Solution to pivoting and implicit NA](https://www.loom.com/share/d49bb263d8194e60b58465a45d1b49c6?sharedAppSource=personal_library)

[Part 2 Separate, Unite and Factors](https://www.loom.com/share/26b5f4b094cd4c8c8bc82c381c71730a)
